---
title: "Lab 1"
author: "Deanna Elliott"
date: "1/6/2022"
output: html_document
---

```{r setup, include=TRUE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(here)
library(sf)
library(tmap)
```

### Read In Data

```{r}

sf_trees <- read_csv(here("data", "sf_trees", "sf_trees.csv"), show_col_types = FALSE)
```

### Wrangling & ggplot Review

**Example 1**: Find counts of observations by legal status and wrangle a bit

```{r}

# Method 1: group_by() %>%  summarize ()

sf_trees %>% 
  group_by(legal_status) %>% 
  summarize(tree_count = n())  # more transparent

# Method 2: different + a few new functions

top_5_status <- sf_trees %>% 
  count(legal_status) %>% 
  drop_na(legal_status) %>% 
  rename(tree_count = n) %>% 
  relocate(tree_count) %>% 
  slice_max(tree_count, n = 5) %>% 
  arrange(-tree_count)    # descending order
```

Make a graph of top 5

```{r}
ggplot(data = top_5_status,
        aes(x = fct_reorder(legal_status, tree_count), y = tree_count)) +
  geom_col(fill = "darkgreen") +
  labs(x = "Legal Status", y = "Tree Count") +
  coord_flip() +
  theme_minimal()
```

**Example 2**: Only keep obs where legal status is "permitted site" and caretaker is "MTA" and store as 'permitted_data_df'

```{r}

sf_trees$legal_status %>%  unique()
unique(sf_trees$caretaker)

permitted_data_df <- sf_trees %>% 
  filter(legal_status == "Permitted Site", 
         caretaker == "MTA")

# filter(legal_status %in% c("Permitted Site", "Private"))
```

**Example 3**: Only keep Blackwood Acacia trees with columns 'legal_status', 'date', 'latitude', 'longitude' and store as 'blackwood_acacia_df'

```{r}

blackwood_acacia_df <- sf_trees %>% 
  filter(str_detect(species, "Blackwood Acacia")) %>% 
  select(legal_status, date, lat = latitude, lon = longitude)

# Make a little graph of locations
ggplot(data = blackwood_acacia_df,
       aes(x = lon, y = lat)) +
  geom_point(color = "darkgreen")
```

**Example 4**: Use 'tidyr::separate()

```{r}

sf_trees_sep <-sf_trees %>% 
  separate(species, into = c("spp_scientific", "spp_common"), sep = " :: ")
```

**Example 5**: Use 'tidyr::unite()

```{r}

ex_5 <- sf_trees %>% 
  unite("id_status", tree_id, legal_status, sep = "_COOL_")
```


### Part 2: Make Some Maps!

**Step 1**: Convert lat/lon to spacial points, 'st_as_sf()'

```{r}

blackwood_acacia_sf <- blackwood_acacia_df %>% 
  drop_na(lat, lon) %>% 
  st_as_sf(coords = c("lon", "lat"))

# We need to tell R what the coord ref system is
st_crs(blackwood_acacia_sf) <- 4326  #means lat/lon system

ggplot(data = blackwood_acacia_sf) +
  geom_sf(color = "darkgreen") +
  theme_minimal()
```

Read in the SF shapefile and add to map

```{r}

sf_map <- read_sf(here("data", "sf_map", "tl_2017_06075_roads.shp"))

sf_map_transform <- st_transform(sf_map, 4326)

ggplot(data = sf_map_transform) +
  geom_sf()
```

Combine the maps!

```{r}

ggplot() +
  geom_sf(data = sf_map,
          size = .1,
          color = "darkgrey") +
  geom_sf(data = blackwood_acacia_sf, 
          color = "red",
          size = 0.5) +
  theme_void() +
  labs(title = "Blackwood Acacias in SF")
```

Now an interactive map

```{r}

tmap_mode("view")

tm_shape(blackwood_acacia_sf) +
  tm_dots()
```

